Optimal (Multiway) Spatial Joins

### 2024-06-03 msj.tex
yufei

\begin{abstract}
    In a {\em spatial join}, we are given a constant number $k \geq 2$ of sets containing axis-parallel rectangles in a 2D space, denoted as $R_1, R_2, ..., R_k$. The objective is to report all $k$-tuples $(r_1, r_2, ..., r_k) \in R_1 \times R_2 \times ... \times R_k$ where the rectangles $r_1, r_2, ..., r_k$ have a non-empty intersection, i.e., $r_1 \cap r_2 \cap ... \cap r_k \neq \emptyset$. The problem holds significant importance in spatial databases and has been extensively studied in the database community. In this paper, we show how to settle the problem in $O(n \log n + \out)$ time --- regardless of the constant $k$ --- where $n = \sum_{i=1}^k |R_i|$ and $\out$ is the result size (i.e., the total number of $k$-tuples reported). The runtime  is asymptotically optimal in the class of comparison-based algorithms, to which our solution belongs. Our result significantly improves the state of the art, which is an algorithm with running time $O(n \log^{2k} n + \out)$.
\end{abstract}

@@

\extraspacing {\bf Math Conventions.} For any integer $x \ge 1$, we use $[x]$ to represent the set $\set{1, 2, ..., x}$. Given $k \ge 2$ sets $S_1, S_2, ..., S_k$ (of arbitrary elements), we often treat a $k$-tuple $(e_1, e_2, .., e_k)$ in the cartesian product $S_1 \times S_2 \times ... \times S_k$ as a $k$-dimensional vector $\bm{t}$ with $\bm{t}[i] = e_i$ for each $i \in [k]$. Unless otherwise specified, every mention of the word ``rectangle'' henceforth will refer to an axis-parallel rectangle in $\real^2$. All logarithms have base 2 by default.

@@

In \cite{kcko22}, Khamis et al.\ investigated a type of joins that extends the conventional equi-join in two ways. First, each attribute value in a relation is an interval (rather than a real value); second, each equality predicate in equi-join is replaced with a ``non-empty intersection'' predicate on the attributes involved. The $k$-SJ problem can be converted to solving a join defined next under the framework of \cite{kcko22}. For each $i \in [k]$, define $R_i$ as a relation over two attributes $X$ and $Y$. For each tuple $\bm{t} \in R_i$, its values $\bm{t}(X)$ and $\bm{t}(Y)$ on the two attributes are both intervals (effectively defining a rectangle). The objective is to output all $k$-tuples $(\bm{t}_1, \bm{t}_2, ..., \bm{t}_k) \in R_1 \times R_2 \times ... \times R_k$ satisfying $\bigcap_{i=1}^k \bm{t}_i(X) \ne \emptyset$ and $\bigcap_{i=1}^k \bm{t}_i(Y) \ne \emptyset$. It is clear that there is one-one correspondence between the result of this join and that of k-SJ. Khamis et al.\ \cite{kcko22} developed an algorithm that can process the join  in $O(n \log^{2k} n + \out)$ time.

@@

It is worth noting that $\Omega(n \log n)$ is a lower bound on the runtime of any comparison-based algorithm  solving the $k$-SJ problem (see Appendix~\ref{app:comp-class} for a formal definition of such algorithms), even for $k = 2$. This can be established via a reduction from the {\em element distinctness} problem; see \cite{dl79}.

@@

\extraspacing {\underline{\em Example 4.1.}} To illustrate our algorithm, we will utilize the running example in Figure~\ref{fig:hv:type1:ex}a, where $k = 4$, and $R_1 = \set{\alpha}$, $R_2 = \set{\beta_2, \beta_2}$, $H$ $=$ $\set{h_1,$ $h_2,$ $..., h_6}$, and $V = \set{v_1, v_2, ..., v_5}$. The set $\J_1$ contains the following tuples: $(\alpha, \beta_2, h_2, v_3)$, $(\alpha, \beta_2, h_2, v_5)$, $(\alpha, \beta_2, h_5, v_3)$, and $(\alpha,$ $\beta_2, h_5, v_6)$. \qed

@@

\extraspacing {\underline{\em Example 4.2.}} Figure~\ref{fig:hv:type1:ex}b shows the rectangles in $R'_1 = \set{\alpha'}$ and $R'_2$ $=$ $\{\beta'_1,$ $\beta'_2\}$. For instance, $\alpha'$, which is trimmed from rectangle $\alpha$, is decided by $h_\bot$ $=$ $h_2$, $h_\top$ $=$ $h_6$, $v_\vdash = v_3$, and $v_\dashv = v_5$. The $(k-2)$-SJ on $R'_1$ and $R'_2$ returns $\J(R_1', R'_2, ..., R'_{k-2}) = \set{(\alpha', \beta_1'), (\alpha', \beta_2')}$.
\qed

@@

\noindent {\underline{\em Example 4.3.} Let us examine, in turn, the two 2-tuples $\bm{t}_1 = (\alpha', \beta_1')$ and $\bm{t}_2 = (\alpha', \beta_2')$ in $\J(R_1', R_2')$. For $\bm{t}_1$, $B_{\bm{t}_1}$ is the rectangle \ttt{ABCD} in Figure~\ref{fig:hv:type1:ex}c, and $\gleft(\bm{t}_1) = \gbot(\bm{t}_1) = \alpha'$. Accordingly, $\dcross_H(\bm{t}_1) = \set{h_2}$ and $\dcross_V(\bm{t}_1) = \set{v_3}$. For $\bm{t}_2$, $B_{\bm{t}_2}$ is the rectangle \ttt{AEFG}, and $\gleft(\bm{t}_2) = \gbot(\bm{t}_2) = \alpha'$. Accordingly, $\dcross_H(\bm{t}_2) = \set{h_2, h_4, h_5}$ and $\dcross_V(\bm{t}_2) = \set{v_3, v_5}$.
\qed

@@

\extraspacing {\underline{\em Example 4.4.} In our running example, the $\J^*$ computed includes 7 tuples: $(\alpha, \beta_1, h_2, v_3)$, and $\{\alpha\} \times \{\beta_2\} \times \set{h_2, h_4, h_5} \times \set{v_3, v_5}$. All these 7 tuples belong to $\J(R_1, R_2, H, V)$ and include the 4 tuples in $\J_1$ (see Example 4.1).
\qed

@@

\extraspacing {\underline{\em Example 4.5.} Recall from Example 4.3 that rectangle $\alpha'$ is both $\gbot(\bm{t}_1)$ and $\gbot(\bm{t}_2)$. Thus, $\maxtop(\alpha') = \max\{\ytop(B_{\bm{t}_1}),$ $\ytop(B_{\bm{t}_2})\} = \max\{6, 9.5\} = 9.5$. The top-sliced rectangle of $\alpha'$ is the rectangle $\alpha^*$ in Figure~\ref{fig:hv:type1:ex}d. Rectangles $\beta'_1$ and $\beta'_2$ do not have top-sliced rectangles.
\qed

@@

\extraspacing {\underline{\em Example 4.6.}} It is clear from Figure~\ref{fig:hv:type1:ex}d that $\cross_H(\alpha^*)$ contains $h_2, h_4$, and $h_5$, sorted in ascending order of y-coordinate. Recall that 
$\gbot(\bm{t}_1) = \gbot(\bm{t}_2) = \alpha'$. Both $\dcross_H(\bm{t}_1) = \set{h_2}$ and $\dcross_H(\bm{t}_2) = \set{h_2, h_4, h_5}$ are prefixes of the sorted list, as is consistent with Lemma~\ref{lmm:hv:type1:cross-r*}.
\qed

@@

\extraspacing {\underline{\em Example 5.1.}} To illustrate our algorithm, we will utilize the running example in Figure~\ref{fig:hv:type2:ex}a, where $k = 4$, and $R_1 = \set{\alpha}$, $R_2 = \set{\beta_2, \beta_2, \beta_3}$, $H$ $=$ $\set{h_1,$ $h_2,$ $h_3}$, and $V = \set{v_1, v_2}$. The set $\J_2$ contains the following tuples: $(\alpha, \beta_1, h_2, v_1)$, $(\alpha, \beta_1, h_3, v_1)$, $(\alpha, \beta_2, h_3, v_1)$, $(\alpha, \beta_3, h_3, v_1)$, and $(\alpha, \beta_3, h_3, v_2)$. \qed

@@

\noindent {\underline{\em Example 5.2.}} Segment $h_1$ has no left-covering rectangle in $R_2$ (see Figure~\ref{fig:hv:type2:ex}a) and thus has no trimmed segment. Segment $h_2$ has one left-covering rectangle in $R_2$, which is $\beta_1$. As the entire $h_2$ is covered by $\beta_1$, it is equivalent to its trimmed segment $h_2'$; see Figure~\ref{fig:hv:type2:ex}b. Segment $h_3$ has two left-covering rectangles in $R_2$, which are $\beta_2$ and $\beta_3$. The right endpoint of its trimmed segment $h_3'$, as shown in Figure~\ref{fig:hv:type2:ex}b, is decided by the right edge of $\beta_3$. Therefore, $H' = \set{h_2', h_3'}$. It is clear from Figure~\ref{fig:hv:type2:ex}b that $\J(R_1, H', V)$ has 3 tuples: $\bm{t}_1 = (\alpha, h_2', v_1), \bm{t}_2 = (\alpha, h_3', v_1)$, and $\bm{t}_3 = (\alpha, h_3', v_2)$.
\qed

@@

\extraspacing {\underline{\em Example 5.3.}} Consider the tuples $\bm{t}_1$, $\bm{t}_2$, and $\bm{t}_3$ of $ \J(R_1, H', V)$ given in Example 5.2. For $\bm{t}_1 = (\alpha, h_2', v_1)$, its effective horizontal segment is $\texttt{DC}$ (see Figure~\ref{fig:hv:type2:ex}b). For $\bm{t}_2 = (\alpha, h_3', v_1)$, its effective horizontal segment is \texttt{EB}. For $\bm{t}_3 = (\alpha, h_3', v_2) \in \J(R_1, H', V)$, its effective horizontal segment is \texttt{EA}. Accordingly, as can be seen from Figure~\ref{fig:hv:type2:ex}c,
$\contained_{R_{k-2}}(\bm{t}_1) = \contained_{R_2}(\bm{t}_1) = \set{\beta_1}$, $\contained_{R_2}(\bm{t}_2) = \set{\beta_1, \beta_2, \beta_3}$, and $\contained_{R_2}(\bm{t}_3) = \set{\beta_3}$.
\qed

@@

\extraspacing {\underline{\em Example 5.4.}} For $\bm{t}_1 = (\alpha, h_2', v_1)$, first find the full segment of $h_2'$, which is $h_2$. As $\beta_1$ is the only rectangle in $\contained_{R_2}(\bm{t}_1)$, Line 5 of the algorithm adds tuple $(\alpha, \beta_1, h_2, v_1)$ to $\J_2$. For $\bm{t}_2 = (\alpha, h_3', v_1)$, the full segment of $h_3'$ is $h_3$. As $\contained_{R_2}(\bm{t}_2) = \set{\beta_1, \beta_2, \beta_3}$, Line 5 adds $(\alpha, \beta_1, h_3, v_1)$, $(\alpha, \beta_2, h_3, v_1)$, and $(\alpha, \beta_3, h_3, v_1)$ to $\J_2$. Finally, the processing of $\bm{t}_3 = (\alpha, h_3', v_2)$ adds $(\alpha, \beta_3, h_3, v_2)$ to $\J_2$.
\qed

@@

\extraspacing {\underline{\em Example 5.5.}} As mentioned, $\J(R_1, H', V)$ has 3 tuples $\bm{t}_1$, $\bm{t}_2$, and $\bm{t}_3$. Both $\bm{t}_2 = (\alpha, h_3', v_1)$ and $\bm{t}_3 = (\alpha, h_3', v_2)$ have $h_3'$ as the horizontal segment. Therefore, $\minleft(h_3')$ equals the smaller between the x-coordinate of $v_1$ and that of $v_2$. The minimal segment $h_3^*$ of $h_3'$ is shown in Figure~\ref{fig:hv:type2:ex}(c). On the other hand, it is easy to verify that $\minleft(h_2')$ is the x-coordinate of $v_1$. The minimal segment $h_2^*$ of $h_2'$ is also shown in Figure~\ref{fig:hv:type2:ex}(c).
\qed

@@

\extraspacing {\underline{\em Example 5.6.}} It is clear from Figure~\ref{fig:hv:type2:ex}(d) that $\contained_{R_2}(h_3^*)$ has rectangles $\beta_3, \beta_2$, $\beta_1$, sorted in descending order of their right boundaries' x-coordinates. Consider $\bm{t}_2 = (\alpha, h_3', v_1)$ and $\bm{t}_3 = (\alpha, h_3', v_2)$. Segment $h_3^*$ is the minimal segment of $h_3'$. As stated in Lemma~\ref{lmm:hv:type2:contained-h*}, both $\contained_{R_2}(\bm{t}_2) = \set{\beta_1, \beta_2, \beta_3}$ and $\contained_{R_2}(\bm{t}_3) = \set{\beta_1}$ are prefixes of the sorted order of $\contained_{R_2}(h_3^*)$. Regarding $h_2^*$, it is the minimal segment of $h_2'$, and $\contained_{R_2}(h_2^*)$ contains only  $\beta_1$. For $\bm{t}_1 = (\alpha, h_2', v_1)$, $\contained_{R_2}(\bm{t}_1) = \set{\beta_1}$ is a (trivial) prefix of $\contained_{R_2}(h_2^*)$, as is consistent with the lemma.
\qed

@@

\def\vgap{\vspace{0.5mm}}
\def\extraspacing{\vspace{1.5mm} \noindent}

>>>>

\begin{abstract}
    In a {\em spatial join}, we are given a constant number $k \geq 2$ of sets --- denoted as $R_1, R_2, ..., R_k$ --- containing axis-parallel rectangles in a 2D space. The objective is to report all $k$-tuples $(r_1, r_2, ..., r_k) \in R_1 \times R_2 \times ... \times R_k$ where the rectangles $r_1, r_2, ..., r_k$ have a non-empty intersection, i.e., $r_1 \cap r_2 \cap ... \cap r_k \neq \emptyset$. The problem holds significant importance in spatial databases and has been extensively studied in the database community. In this paper, we show how to settle the problem in $O(n \log n + \out)$ time --- regardless of the constant $k$ --- where $n = \sum_{i=1}^k |R_i|$ and $\out$ is the result size (i.e., the total number of $k$-tuples reported). The runtime  is asymptotically optimal in the class of comparison-based algorithms, to which our solution belongs. Previously, the state of the art was an algorithm with running time $O(n \log^{2k} n + \out)$.
\end{abstract}

@@

\extraspacing {\bf Math Conventions.} For any integer $x \ge 1$, we use $[x]$ to represent the set $\set{1, 2, ..., x}$. Given $k \ge 2$ sets $S_1, S_2, ..., S_k$ (of arbitrary elements), we often treat a $k$-tuple $(e_1, e_2, .., e_k)$ in the cartesian product $S_1 \times S_2 \times ... \times S_k$ as a $k$-dimensional vector $\bm{t}$ with $\bm{t}[i] = e_i$ for each $i \in [k]$. Unless otherwise stated, every mention of the word ``rectangle'' henceforth will refer to an axis-parallel rectangle in $\real^2$. All logarithms have base 2 by default.

@@

In \cite{kcko22}, Khamis et al.\ investigated a type of joins that extends the conventional equi-join in two ways. First, each attribute value in a relation is an interval (rather than a real value); second, each equality predicate in equi-join is replaced with a ``non-empty intersection'' predicate on the attributes involved. The $k$-SJ problem can be converted to a join under the framework of \cite{kcko22} as defined next. For each $i \in [k]$, define $R_i$ as a relation over two attributes $X$ and $Y$. For each tuple $\bm{t} \in R_i$, its values $\bm{t}(X)$ and $\bm{t}(Y)$ on the two attributes are both intervals (effectively defining a rectangle). The objective is to output all $k$-tuples $(\bm{t}_1, \bm{t}_2, ..., \bm{t}_k) \in R_1 \times R_2 \times ... \times R_k$ satisfying $\bigcap_{i=1}^k \bm{t}_i(X) \ne \emptyset$ and $\bigcap_{i=1}^k \bm{t}_i(Y) \ne \emptyset$. It is clear that there is one-one correspondence between the result of this join and that of k-SJ. Khamis et al.\ \cite{kcko22} developed an algorithm that can process the join  in $O(n \log^{2k} n + \out)$ time.

@@

It is worth noting that $\Omega(n \log n)$ is a lower bound on the runtime of any comparison-based $k$-SJ algorithms (see Appendix~\ref{app:comp-class} for a formal definition of such algorithms), even for $k = 2$. This can be established via a reduction from the {\em element distinctness} problem; see \cite{dl79}.

@@

\extraspacing {\underline{\em Example 4.1.}} To illustrate our algorithm, we will utilize the running example in Figure~\ref{fig:hv:type1:ex}a, where $k = 4$, and $R_1 = \set{\alpha}$, $R_2 = \set{\beta_2, \beta_2}$, $H$ $=$ $\set{h_1,$ $h_2,$ $..., h_6}$, and $V = \set{v_1, v_2, ..., v_5}$. The set $\J_1$ contains the following tuples: $(\alpha, \beta_2, h_2, v_3)$, $(\alpha, \beta_2, h_2, v_5)$, $(\alpha, \beta_2, h_5, v_3)$, and $(\alpha,$ $\beta_2, h_5, v_6)$. \done

@@

\extraspacing {\underline{\em Example 4.2.}} Figure~\ref{fig:hv:type1:ex}b shows the rectangles in $R'_1 = \set{\alpha'}$ and $R'_2$ $=$ $\{\beta'_1,$ $\beta'_2\}$. For instance, $\alpha'$, which is trimmed from rectangle $\alpha$, is decided by $h_\bot$ $=$ $h_2$, $h_\top$ $=$ $h_6$, $v_\vdash = v_3$, and $v_\dashv = v_5$. The $(k-2)$-SJ on $R'_1$ and $R'_2$ returns $\J(R_1', R'_2, ..., R'_{k-2}) = \set{(\alpha', \beta_1'), (\alpha', \beta_2')}$.
\done

@@

\noindent {\underline{\em Example 4.3.} Let us examine, in turn, the two 2-tuples $\bm{t}_1 = (\alpha', \beta_1')$ and $\bm{t}_2 = (\alpha', \beta_2')$ in $\J(R_1', R_2')$. For $\bm{t}_1$, $B_{\bm{t}_1}$ is the rectangle \ttt{ABCD} in Figure~\ref{fig:hv:type1:ex}c, and $\gleft(\bm{t}_1) = \gbot(\bm{t}_1) = \alpha'$. Accordingly, $\dcross_H(\bm{t}_1) = \set{h_2}$ and $\dcross_V(\bm{t}_1) = \set{v_3}$. For $\bm{t}_2$, $B_{\bm{t}_2}$ is the rectangle \ttt{AEFG}, and $\gleft(\bm{t}_2) = \gbot(\bm{t}_2) = \alpha'$. Accordingly, $\dcross_H(\bm{t}_2) = \set{h_2, h_4, h_5}$ and $\dcross_V(\bm{t}_2) = \set{v_3, v_5}$.
\done

@@

\extraspacing {\underline{\em Example 4.4.} In our running example, the $\J^*$ computed includes 7 tuples: $(\alpha, \beta_1, h_2, v_3)$, and $\{\alpha\} \times \{\beta_2\} \times \set{h_2, h_4, h_5} \times \set{v_3, v_5}$. All these 7 tuples belong to $\J(R_1, R_2, H, V)$ and include the 4 tuples in $\J_1$ (see Example 4.1).
\done

@@

\extraspacing {\underline{\em Example 4.5.} Recall from Example 4.3 that rectangle $\alpha'$ is both $\gbot(\bm{t}_1)$ and $\gbot(\bm{t}_2)$. Thus, $\maxtop(\alpha') = \max\{\ytop(B_{\bm{t}_1}),$ $\ytop(B_{\bm{t}_2})\} = \max\{6, 9.5\} = 9.5$. The top-sliced rectangle of $\alpha'$ is the rectangle $\alpha^*$ in Figure~\ref{fig:hv:type1:ex}d. Rectangles $\beta'_1$ and $\beta'_2$ do not have top-sliced rectangles.
\done

@@

\extraspacing {\underline{\em Example 4.6.}} It is clear from Figure~\ref{fig:hv:type1:ex}d that $\cross_H(\alpha^*)$ contains $h_2, h_4$, and $h_5$, sorted in ascending order of y-coordinate. Recall that 
$\gbot(\bm{t}_1) = \gbot(\bm{t}_2) = \alpha'$. Both $\dcross_H(\bm{t}_1) = \set{h_2}$ and $\dcross_H(\bm{t}_2) = \set{h_2, h_4, h_5}$ are prefixes of the sorted list, as is consistent with Lemma~\ref{lmm:hv:type1:cross-r*}.
\done

@@

\extraspacing {\underline{\em Example 5.1.}} To illustrate our algorithm, we will utilize the running example in Figure~\ref{fig:hv:type2:ex}a, where $k = 4$, and $R_1 = \set{\alpha}$, $R_2 = \set{\beta_2, \beta_2, \beta_3}$, $H$ $=$ $\set{h_1,$ $h_2,$ $h_3}$, and $V = \set{v_1, v_2}$. The set $\J_2$ contains the following tuples: $(\alpha, \beta_1, h_2, v_1)$, $(\alpha, \beta_1, h_3, v_1)$, $(\alpha, \beta_2, h_3, v_1)$, $(\alpha, \beta_3, h_3, v_1)$, and $(\alpha, \beta_3, h_3, v_2)$. \done

@@

\noindent {\underline{\em Example 5.2.}} Segment $h_1$ has no left-covering rectangle in $R_2$ (see Figure~\ref{fig:hv:type2:ex}a) and thus has no trimmed segment. Segment $h_2$ has one left-covering rectangle in $R_2$, which is $\beta_1$. As the entire $h_2$ is covered by $\beta_1$, it is equivalent to its trimmed segment $h_2'$; see Figure~\ref{fig:hv:type2:ex}b. Segment $h_3$ has two left-covering rectangles in $R_2$, which are $\beta_2$ and $\beta_3$. The right endpoint of its trimmed segment $h_3'$, as shown in Figure~\ref{fig:hv:type2:ex}b, is decided by the right edge of $\beta_3$. Therefore, $H' = \set{h_2', h_3'}$. It is clear from Figure~\ref{fig:hv:type2:ex}b that $\J(R_1, H', V)$ has 3 tuples: $\bm{t}_1 = (\alpha, h_2', v_1), \bm{t}_2 = (\alpha, h_3', v_1)$, and $\bm{t}_3 = (\alpha, h_3', v_2)$.
\done

@@

\extraspacing {\underline{\em Example 5.3.}} Consider the tuples $\bm{t}_1$, $\bm{t}_2$, and $\bm{t}_3$ of $ \J(R_1, H', V)$ given in Example 5.2. For $\bm{t}_1 = (\alpha, h_2', v_1)$, its effective horizontal segment is $\texttt{DC}$ (see Figure~\ref{fig:hv:type2:ex}b). For $\bm{t}_2 = (\alpha, h_3', v_1)$, its effective horizontal segment is \texttt{EB}. For $\bm{t}_3 = (\alpha, h_3', v_2) \in \J(R_1, H', V)$, its effective horizontal segment is \texttt{EA}. Accordingly, as can be seen from Figure~\ref{fig:hv:type2:ex}c,
$\contained_{R_{k-2}}(\bm{t}_1) = \contained_{R_2}(\bm{t}_1) = \set{\beta_1}$, $\contained_{R_2}(\bm{t}_2) = \set{\beta_1, \beta_2, \beta_3}$, and $\contained_{R_2}(\bm{t}_3) = \set{\beta_3}$.
\done

@@

\extraspacing {\underline{\em Example 5.4.}} For $\bm{t}_1 = (\alpha, h_2', v_1)$, first find the full segment of $h_2'$, which is $h_2$. As $\beta_1$ is the only rectangle in $\contained_{R_2}(\bm{t}_1)$, Line 5 of the algorithm adds tuple $(\alpha, \beta_1, h_2, v_1)$ to $\J_2$. For $\bm{t}_2 = (\alpha, h_3', v_1)$, the full segment of $h_3'$ is $h_3$. As $\contained_{R_2}(\bm{t}_2) = \set{\beta_1, \beta_2, \beta_3}$, Line 5 adds $(\alpha, \beta_1, h_3, v_1)$, $(\alpha, \beta_2, h_3, v_1)$, and $(\alpha, \beta_3, h_3, v_1)$ to $\J_2$. Finally, the processing of $\bm{t}_3 = (\alpha, h_3', v_2)$ adds $(\alpha, \beta_3, h_3, v_2)$ to $\J_2$.
\done

@@

\extraspacing {\underline{\em Example 5.5.}} As mentioned, $\J(R_1, H', V)$ has 3 tuples $\bm{t}_1$, $\bm{t}_2$, and $\bm{t}_3$. Both $\bm{t}_2 = (\alpha, h_3', v_1)$ and $\bm{t}_3 = (\alpha, h_3', v_2)$ have $h_3'$ as the horizontal segment. Therefore, $\minleft(h_3')$ equals the smaller between the x-coordinate of $v_1$ and that of $v_2$. The minimal segment $h_3^*$ of $h_3'$ is shown in Figure~\ref{fig:hv:type2:ex}(c). On the other hand, it is easy to verify that $\minleft(h_2')$ is the x-coordinate of $v_1$. The minimal segment $h_2^*$ of $h_2'$ is also shown in Figure~\ref{fig:hv:type2:ex}(c).
\done

@@

\extraspacing {\underline{\em Example 5.6.}} It is clear from Figure~\ref{fig:hv:type2:ex}(d) that $\contained_{R_2}(h_3^*)$ has rectangles $\beta_3, \beta_2$, $\beta_1$, sorted in descending order of their right boundaries' x-coordinates. Consider $\bm{t}_2 = (\alpha, h_3', v_1)$ and $\bm{t}_3 = (\alpha, h_3', v_2)$. Segment $h_3^*$ is the minimal segment of $h_3'$. As stated in Lemma~\ref{lmm:hv:type2:contained-h*}, both $\contained_{R_2}(\bm{t}_2) = \set{\beta_1, \beta_2, \beta_3}$ and $\contained_{R_2}(\bm{t}_3) = \set{\beta_1}$ are prefixes of the sorted order of $\contained_{R_2}(h_3^*)$. Regarding $h_2^*$, it is the minimal segment of $h_2'$, and $\contained_{R_2}(h_2^*)$ contains only  $\beta_1$. For $\bm{t}_1 = (\alpha, h_2', v_1)$, $\contained_{R_2}(\bm{t}_1) = \set{\beta_1}$ is a (trivial) prefix of $\contained_{R_2}(h_2^*)$, as is consistent with the lemma.
\done

@@

% \def\vgap{\vspace{0.5mm}}
% \def\extraspacing{\vspace{1.5mm} \noindent}
